Combinatorics of the geometry of Wilson loop diagrams II: Grassmann necklaces, dimensions, and denominators

### 2019-10-06 GWLDCII.tex
minor grammar + notation updates

In the next two lemmas, we consider the plusses contributed to the Le diagram of $W$ by $I_{n-2}^W$.

@@

In this section, we investigate the poles of the integrand for $I(W)$, that is $R(W)$ as defined in Definition~\ref{def R(W)}. Since SYM $N=4$ is a finite theory \cite{Alday:2007hr, ParkeTaylorids}, the scattering amplitudes (and thus the holomorphic Wilson loops $\,\mathfrak{W}_{k,n}$) are finite. That is, for a fixed set of external particles $\cZ$, \bas \sum_{W \text{ s.t. } |\cP| = k} I(W)(\cZ) <\infty \;.\eas In otherwords, all poles of the intergrands must cancel. As a step towards this, in this section, we prove some results about the structure of the denominator of the integrand, $R(W)$.

The results of Section~\ref{sec GN algorithm} allow us to relate the configuration of propagators in a Wilson loop diagram $W$ to minors of $C(W)$, $R(W)$.  In this section fix an arbitrary ordering of the propagators, equivalently the rows of $C(W)$.

@@

\begin{algorithm}\label{alg WLD to denom via GN}
Let $W = (\cP,[n])$ be a Wilson loop diagram, and let $C(W)$ be the matrix of $W$ as defined in \eqref{C(W) dfn} with a fixed ordering on $\cP$. (see Section~\ref{section WLD background}).
\begin{itemize}
  \item For each $i \in [n]$, we construct a factor $r_i$ as follows:
    \begin{itemize}
      %\item Use Algorithm~\ref{alg:put GN on WLD} to obtain a bijection between the propagators and the $i-1$st Grassmann necklace element, $I_{i-1}$.  (By convention set $I_{0}=I_n$.)  Write $I_{i-1}(p)$ for the vertex associated to propagator $p$ under this bijection.
      %\item Use Algorithm~\ref{alg:put GN on WLD} to obtain a bijection between the propagators and the $i$th Grassmann necklace element, $I_{i}$.  Write $I_{i}(p)$ for the vertex associated to propagator $p$ under this bijection.
      \item Let $S_i = \{p \in \cP \ | \ I_{i-1}(p) \neq I_i(p)\}$. (By convention, set $I_{0} = I_n$.)
	  %\item Write $\Delta_{I_i}$ for the determinant of the $k \times k$ minor of $C(W)$ with columns indexed by $I_i$.
      \item Let $r_i$ be the determinant of the $|S_i|\times |S_i|$ minor of $C(W)$ with rows indexed by $S_i$ and columns indexed by $I_i(S_i)$.  If $S_i = \emptyset$, set $r_i = 1$. %$\Delta_{I_i}$ with all variables from rows associated to $p\not\in S_i$ set to $1$.
    \end{itemize}
  \item Define $R = \prod_{i=1}^n r_i$.
\end{itemize}
\end{algorithm}

That is, each $r_i$ is the determinant of the minor determined by the propagators (rows) that contribute different values (columns) to the $i^{th}$ element of the Grassmann necklace than they did to the $i-1^{st}$.

As we discuss the algorithm it will also be useful to have the notation $\Delta_{I_i}$ for the determinant of the $k\times k$ minor of $C(W)$ with columns indexed by $I_i$. In particular, one summand of this determinant is given by $\prod_{p \in \cP}x_{p, I_i(p)}$. The other summands are (up to sign) given by other bijections between the propagator set $\cP$ and the vertex set $I_i$.

As we take each $r_i$ to simply be the determinant of a particular submatrix of $C(W)$, the sign of each $r_i$ is well-defined (since an ordering on the propagators has been fixed).  Before proving the main results of this section, we given an illustrative example. Below, we show that $R$ is equal to the denominator of the Wilson loop diagram as defined in Definition~\ref{def R(W)}, and is the radical of the product $\prod_{i=1}^n \Delta_{I_i}$.
%  However, we will not be interested in these signs as they only contribute an overall sign to $R$ and our main theorem in this section (Theorem~\ref{thm denom}) will simply be to show that $R$ is equal up to a $\mathbb{Q}$-multiple to the denominator of the Wilson loop diagram (see Definition~\ref{def R(W)}).

@@

\begin{eg}
Consider the Wilson loop diagram in Figure~\ref{fig R eg}. Assigning propagators $p$, $q$, $r$ to rows $1,2,3$ respectively, we obtain the matrix
\[
C(W) = \begin{bmatrix} a & b & 0 & 0 & 0 & c & d \\ e & f & 0 & 0 & g & h & 0 \\ i & j & 0 & k & l & 0 & 0 \end{bmatrix}
\]
The Grassmann necklace of this diagram is 
\begin{gather*}I_1 = \{1,2,4\}, I_2 = \{2,4,5\}, I_3 = \{4,5,6\}, I_4=\{4,5,6\},\\ I_5=\{5,6,7\}, I_6 = \{6,7,1\}, I_7=\{7,1,2\}. \end{gather*}  
Figure~\ref{fig R eg} indicates indicates which vertex is contributed to $I_i$ by each propagator for each $i \in [1,7]$.

From $I_1$ to $I_2$, the propagators $p$ and $q$ change which vertices they contribute. That is $I_1(p) = 1$, $I_1(q)= 2$ while $I_2(p) = 2$, $I_2(q)= 5$. However, $r$ contributes  4 to both $I_1$ and $I_2$, so $S_2 = \{p,q\}$.  Then
\[
\Delta_{I_2}=\det\begin{bmatrix} b & 0 & 0 \\ f & 0 & g \\ j & k & l \end{bmatrix} = kgb, \qquad r_2 = \det\begin{bmatrix} b & 0 \\ f & g \end{bmatrix} = gb.
\] 
Continuing likewise, we get $r_3 = c$, $r_4=1$ (since $I_4 = I_3$), $r_5 = lhd$, and $r_6 = i$.

At $I_7$ we see for the first time an irreducible quadratic factor: we have $S_7 = \{q,s\}$ with $I_7(S_7) = \{ 1,2\}$, $\Delta_{I_7} = d(ej-fi)$ and $r_7 = ej-fi$.

@@

\begin{proof}
\begin{enumerate}
    \item The nonzero entries of $C(W)$ are independent indeterminates and so every $i\times i$ minor of $C(W)$ is either homogeneous of degree $i$ or is $0$.  Thus each $\Delta_{I_i}$ and each $r_i$ is homogeneous.  %Furthermore, each row contributes one factor to each term in the expansion of $\Delta_{I_i}$ so the result of setting the variables from a subset of rows to $1$ is still homogeneous.  Thus each $r_i$ is homogeneous.
    \item Using the expression for the determinant as a sum over permutations along with Corollary~\ref{lem basis as perm} we see that $\Delta_{I_i}$ is a sum over bijections between $I_i$ and $\mathcal{P}$.  The nonzero terms in this sum are precisely those bijections such that each propagator is associated to one of its supporting vertices in $I_i$, namely $I_i(p)$, since only those locations in $C(W)$ are nonzero.  Since the nonzero entries of $C(W)$ are independent there can be no cancellation between terms in this expansion.

@@

All that remains is to prove that $\Delta_{I_i}$ has no irreducible factors of degree 3 or more.  Suppose for a contradiction that $f$ is a factor of $\Delta_{I_i}$ of degree $\geq 3$. Without loss of generality\footnote{This is without loss of generality for the following reason.  By removing the propagators which come before those contributing to $f$ in the order given by $I_i$ and changing $i$ to be the first vertex which contributes to $f$, we obtain an admissible diagram.  In this diagram $f$ still divides $\Delta_{I_i}$ and also $i\in I_i$ and
$i$ contributes to $f$.  Thus this diagram can be used in place of $W$.} we may assume $i\in I_i$ and $i$ contributes to $f$.
Finally, we can suppose that $W$ is minimal in number of propagators with an irreducible factor of degree $\geq 3$.

Let $p$ be the propagator such that $I_i(p) = i$. There are two cases to consider, when $p$ is supported on the edge $i-1$ or the edge $i$.  These are illustrated in Figure~\ref{fig no big factors}

@@

\textbf{Case 1}: Suppose $p = (i-1, m)$. Then $p$ has one end on $i-1^{th}$ edge and $I_{i+1}(p) = m$ by Remark~\ref{rmk cyclic}.

@@

If neither $m$ nor $m+1$ appear in $I_i$ then by Corollary~\ref{no coloops} $V(p) \cap I_i = \{i\}$, and so the row of $p$ in the matrix of $\Delta_{I_i}$ has only one nonzero entry; hence $\Delta_{I_i}$ has a linear factor contributed by $p$ and $i$, which is a contradiction to the fact that $i$ contributes to $f$.  So we must have at least one of $m$ and $m+1$ in $I_i$.  However, all propagators in $S$ contribute to $I_{i}$ strictly before $m$ by \st{the previous paragraph}{\color{violet}Remark {clockwise ordering rem}}, so permuting rows and columns as needed, the matrix giving $\Delta_{I_i}$ has the form
\[
\begin{bmatrix} A & B \\ 0 & C\end{bmatrix}
\]
where $A$ is the $|S|\times |S|$ matrix indexed by the propagators in $S$ and the vertices in $I_i(S)$. No other propagators can be supported on these vertices since all other propagators are outside of $p$, and $p$ is the first propagator supported at $i$; this explains the zero block.  Therefore $\Delta_{I_i} = \det A \det C$, and both factors are nontrivial since at least one of $m$ and $m+1$ appear in $I_i$.  Since $p\in \cP_{in}(p)$ we have $f|\det A${\color{violet}. If we remove the propagators that contribute variables to the minor $C$,}\st{ and if we remove the propagator outside of $p$ that contributes $m$ or $m+1$,} we get a diagram $W'$ with fewer propagators for which $\det A | \Delta_{I^{W'}_i}$ still holds, hence $f|\Delta_{I^{W'}_i}$.  This contradicts the minimality of our choices. \sanote{trying to make clear that the minor changes with the diagram. May have just made it too notation heavy.}

\textbf{Case 2}: Suppose $p = (i, m)$, i.e. has an end on the $i^{th}$ edge. If no other propagators are supported on the vertex $i$ then the corresponding column of $C(W)$ has only one nonzero entry in it, and so $\Delta_{I_i}$ has a linear factor contributed by $p$ and $i$; as above, this is a contradiction. \st{Since $I_i(p) = i$ there can be no propagator supported on $i$ with an end on the edge $(i-1, i)$.  Thus to have another propagator supported on $i$ there is at least one other propagator on edge $(i,i+1)$.  Further since since $I_i(p)=i$, one of these other propagators contributes $i+1$ in $I_i$, so we can take $q$ to be the propagator such that $I_i(q)=i+1$.  We know that $q$ has one end on the edge $(i, i+1)$ and is adjacent to $p$ on that edge in the counterclockwise direction (see Figure {\ref{fig no big factors}}).}  \hlfix{Since $p$ is the first propagator in the ordering imposed by $I_i$, it is the counterclockwise most propagator supported on $i$ (by Remark {\ref{clockwise ordering rem}}). Since there are multiple propagators supported on $i$, there must be a $q$ ending on edge $i$, such that $I_i(q) = i+1$, that is second clockwisemost propagators supported by $i$.}{Alt. proof} Write $q = (i, j)$ for the support of $q$.  The situation for $q$ is very similar to case 1: \st{in particular, we have $I_{i+1}(q) = j$ by Lemma {\ref{vertex cyclic int lem}} and so if $j\in I_i$ then the propagator which contributes $j$ is outside of $q$.}{\color{violet}if we have $r$ such that $I_i(r) = j$, as above, $r$ lies outside $q$: $r \in \cP_{out}(q, <_i)$. }

Similarly to Case 1, let $S = \cP_{in}(q) \cup \{p\}$ in the $<_i$ order. Then all propagators in $S$ contribute to $I_i$ strictly before $j$ and no other propagators are supported on vertices strictly before $j$.  Thus the matrix giving $\Delta_{I_i}$ has the form {\color{violet}(when reordered according to how they contribute to $I_i$)}
\[
\begin{bmatrix} A & B \\ 0 & C\end{bmatrix}
\]
where $A$ is the submatrix indexed by the propagators in $S$ and the vertices in $I_i(S)$. Again two things can now happen.  If some vertex $j$ or larger (with respect to $>_i$) belongs to $I_i$ then $B$ and $C$ are at least one column wide, and so the block form of the matrix gives a nontrivial factorization of $\Delta_{I_i}=\det A\det B$.  This yields a contradiction as in Case 1: $f|\det A$ and removing a propagator not in $S$ gives a smaller diagram with a factor of degree $\geq 3$ which contradicts our minimality assumption.\note{I rephrased this last sentence quite a bit because earlier things had gotten rephrased} \sanote{reads fine to me.}

On the other hand, if no vertex $\geq_i j$ is in $I_i$ then $\Delta_{I_i} = \det A$.  Looking in more detail into $A$, note that the only vertices supporting both $p$ and $q$ which belong to $I_i$ are $i$ and $i+1$. Hence
\[
A = \begin{bmatrix} D & 0 \\ E & F\end{bmatrix}
\]
where $D$ is the $2\times 2$ matrix indexed by the propagators $p$ and $q$ and the vertices $i$ and $i+1$.  Thus $p$ and $i$ contribute to a quadratic factor of $\Delta_{I_i}$, once again contradicting our assumptions.

@@

\item \sanote{changed $b$ to $a+1$ here.} Suppose propagators $p$ and $q$ are supported on a common edge  $(a, a+1)$, with $I_i(p)=a$ and $I_i(q)=a+1$.  Let $x_{p,a},x_{p,a+1},x_{q,a},x_{q,a+1}$ be the associated variables in $C(W)$. For any fixed bijection $\sigma$ from $\cP-\{p,q\}$ to $I_i -\{a,a+1\}$ for which each propagator is supported on its image under the bijection, we can extend $\sigma$ to a bijection from propagators to $I_i$ in two ways: either $p\mapsto a$ and $q\mapsto a+1$ or $p\mapsto a+1$ and $q\mapsto a$.  The sum of the contributions of all these bijections to $\Delta_{I_i}$ is therefore the product of $x_{p,a}x_{q,a+1}-x_{p,a+1}x_{q,a}$ {\color{violet} By the arguments in the proof of the claim in part 2 of this theorem} \note{Could this use a ref back to where we discuss this kind of arugment in more detail.}\sanote{see violet?} with the minor coming from $\cP-\{p, q\}$ and $I_i - \{a,a+1\}$.  Since there is no cancellation of terms in the expansion of $\Delta_{I_i}$, if any other terms appear then they will cause a factor which is not in the form described in the previous part.  Therefore no such terms exist and $x_{p,a}x_{q,a+1}-x_{p,a+1}x_{q,a}$ is a factor of $\Delta_{I_i}$.

Now let $f$ be a quadratic factor of $\Delta_{I_i}$.  By part (2) we know that $f$ is a $2\times 2$ minor coming from two propagators, call them $p$ and $q$, and two vertices, call them $a <_i b$.  It remains to show that \st{$a$ and $b$ are adjacent} {\color{violet}$b = a+1$}.  From this we can conclude that $p$ and $q$ each have one end on $(a,a+1)$. \st{, as any other way for both $p$ and $q$ to be supported on two consecutive vertices would contradict noncrossing or the density requirement of admissibility.}

@@

\item Consider $p\in S_i$ from \ref{alg WLD to denom via GN}, and note that $\Delta_{I_i}$ is homogeneous linear in the variables of the row corresponding to $p$.  By part (2), either exactly one variable in the row corresponding to $p$ appears in $\Delta_{I_i}$ and this variable is a factor of $\Delta_{I_i}$, or exactly two variables from the row corresponding to $p$ appear in $\Delta_{I_i}$ and they appear as part of a quadratic factor.  In the first case let the variable be $x$. Then $x$ is a factor of $\Delta_{I_i}$. {\color{violet} Therefore, $x$ must correspond to entry in the row of $p$ and column of $I_i(p)$, i.e. $x = x_{p, I_i(p)}$.} \st{and so in particular the monomial in $\Delta_{I_i}$ corresponding to the bijection between propagators and vertices of $I_i$ associates the column of $x$ to $p$. Thus $x$ also appears in $r_i$ and s} Since the matrix for $r_i$ is a minor of the matrix for $\Delta_{I_i}$ \st{and every term in $\Delta_{I_i}$ involves $x$,} we also have that \st{every term in $r_i$ involves $x$ so}{\color{violet}$x$ is a factor of $r_i$.} \st{$x$ is a factor of both $r_i$ and $\Delta_{I_i}$ and is the only variable from this row in either polynomial.}

Now suppose two variables from the row $p$ appear in a quadratic factor $f$ of $\Delta_{I_i}$.  By part (3), there is another propagator $q$ and an edge $(a,a+1)$ such that $f$ is the $2\times 2$ minor coming from $p, q$ and $a, a+1$, with \st{$I_i(p)=a$, $I_i(q)=b$}.  {\color{violet}There are two situations which can occur, either $I_i(p)=a$ and $I_i(q)=a+1$, or $I_i(q)=a$ and $I_i(p)=a+1$. These are} illustrated in Figure~\ref{fig quadratic}; we show that in both cases it follows that $q \in S_i$ as well.

>>>>

In the next two lemmas, we consider the plusses contributed to the Le diagram of $W$ by $I_{n-2}^W$. Lemma~\ref{lem I} lists three possibilities for $I_{n-2}^W$, but the following observation shows that we need only consider two of them.

@@

In this section we investigate the poles of the integrand for $I(W)$. Since SYM $N=4$ is a finite theory \cite{Alday:2007hr, ParkeTaylorids}, the scattering amplitudes (and thus the holomorphic Wilson loops $\,\mathfrak{W}_{k,n}$) are finite. That is, for a fixed set of external particles $\cZ$, we must have
\bas \sum_{W \text{ s.t. } |\cP| = k} I(W)(\cZ) <\infty \;.\eas
In other words, all poles of the intergrands must cancel. As a step towards this, in this section we prove some results about the structure of the denominator $R(W)$ of the integrand, as defined in Definition~\ref{def R(W)}.

The results of Section~\ref{sec GN algorithm} allow us to relate the configuration of propagators in a Wilson loop diagram $W$ to minors of $C(W)$, $R(W)$.  In this section fix an arbitrary ordering of the propagators, equivalently the rows of $C(W)$.\note{There's some grammatical issues in this paragraph and I'm not sure what it's supposed to be saying.}

@@

\begin{algorithm}\label{alg WLD to denom via GN}
Let $W = (\cP,[n])$ be a Wilson loop diagram, and let $C(W)$ be the matrix of $W$ as defined in equation \eqref{C(W) dfn} with a fixed ordering on $\cP$ (see Section~\ref{section WLD background})\note{This was written before the exposition above; now I think we can simplify it to "Let $W$ be a WLD and $C(W)$ its associated matrix?}.
\begin{itemize}
  \item For each $i \in [n]$, we construct a factor $r_i$ as follows:
    \begin{itemize}
      %\item Use Algorithm~\ref{alg:put GN on WLD} to obtain a bijection between the propagators and the $i-1$st Grassmann necklace element, $I_{i-1}$.  (By convention set $I_{0}=I_n$.)  Write $I_{i-1}(p)$ for the vertex associated to propagator $p$ under this bijection.
      %\item Use Algorithm~\ref{alg:put GN on WLD} to obtain a bijection between the propagators and the $i$th Grassmann necklace element, $I_{i}$.  Write $I_{i}(p)$ for the vertex associated to propagator $p$ under this bijection.
      \item Let $S_i = \{p \in \cP \ | \ I_{i-1}(p) \neq I_i(p)\}$. (By convention, set $I_{0} = I_n$.)
	  %\item Write $\Delta_{I_i}$ for the determinant of the $k \times k$ minor of $C(W)$ with columns indexed by $I_i$.
      \item Let $r_i$ be the determinant of the $|S_i|\times |S_i|$ minor of $C(W)$ with rows indexed by $S_i$ and columns indexed by $I_i(S_i)$.  If $S_i = \emptyset$, set $r_i = 1$. %$\Delta_{I_i}$ with all variables from rows associated to $p\not\in S_i$ set to $1$.
    \end{itemize}
  \item Define $R = \prod_{i=1}^n r_i$.
\end{itemize}
\end{algorithm}

That is, each $r_i$ is the determinant of the minor determined by the propagators (rows) that contribute different values (columns) to the $i^{th}$ element of the Grassmann necklace than they did to the $i-1^{st}$. Since each $r_i$ is simply the determinant of a particular submatrix of $C(W)$ and an ordering on the propagators has been fixed, the sign of each $r_i$ is well-defined.

We also introduce the notation $\Delta_{I_i}$ to represent the determinant of the $k\times k$ minor of $C(W)$ with columns indexed by $I_i$. Note that one summand of this determinant is given by $\prod_{p \in \cP}x_{p, I_i(p)}$; the other summands are (up to sign) given by other bijections between the propagator set $\cP$ and the vertex set $I_i$.

Below, we show that the polynomial $R$ associated to Wilson loop diagram $W$ is equal to the denominator $R(W)$ from Definition~\ref{def R(W)}, and is the radical of the product $\prod_{i=1}^n \Delta_{I_i}$. First, however, we give a worked example of Algorithm~\ref{alg WLD to denom via GN}.

@@

\begin{eg}
Consider the Wilson loop diagram in Figure~\ref{fig R eg}. Assigning propagators $p$, $q$, $r$ to rows $1,2,3$ respectively, we obtain the matrix
\[
C(W) = \begin{bmatrix} a & b & 0 & 0 & 0 & c & d \\ e & f & 0 & 0 & g & h & 0 \\ i & j & 0 & k & l & 0 & 0 \end{bmatrix}
\]
The Grassmann necklace of this diagram is 
\begin{gather*}I_1 = \{1,2,4\}, I_2 = \{2,4,5\}, I_3 = \{4,5,6\}, I_4=\{4,5,6\},\\ I_5=\{5,6,7\}, I_6 = \{6,7,1\}, I_7=\{7,1,2\}. \end{gather*}  
Figure~\ref{fig R eg} indicates indicates which vertex is contributed to $I_i$ by each propagator, for each $i \in [1,7]$.

From $I_1$ to $I_2$, the propagators $p$ and $q$ change which vertices they contribute to; specifically, we have $I_1(p) = 1$ and $I_1(q)= 2$, while $I_2(p) = 2$ and $I_2(q)= 5$. On the other hand, $r$ contributes  4 to both $I_1$ and $I_2$. We therefore have $S_2 = \{p,q\}$.  The associated minors $\Delta_{I_2}$ and $r_2$ are therefore
\[
\Delta_{I_2}=\det\begin{bmatrix} b & 0 & 0 \\ f & 0 & g \\ j & k & l \end{bmatrix} = kgb, \qquad r_2 = \det\begin{bmatrix} b & 0 \\ f & g \end{bmatrix} = gb.
\] 
Continuing likewise, we get $r_3 = c$, $r_4=1$ (since $I_4 = I_3$), $r_5 = lhd$, and $r_6 = i$.

At $I_7$ we see for the first time an irreducible quadratic factor: we have $S_7 = \{q,s\}$ with ${I_7(S_7) = \{ 1,2\}}$, $\Delta_{I_7} = d(ej-fi)$ and $r_7 = ej-fi$.

@@

\begin{proof}
\begin{enumerate}
    \item The nonzero entries of $C(W)$ are independent indeterminates and so every $i\times i$ minor of $C(W)$ is either homogeneous of degree $i$ or is $0$.  Thus each $\Delta_{I_i}$ and each $r_i$ is homogeneous.  %Furthermore, each row contributes one factor to each term in the expansion of $\Delta_{I_i}$ so the result of setting the variables from a subset of rows to $1$ is still homogeneous.  Thus each $r_i$ is homogeneous.
    \item Viewing the determinant as a sum over permutations, and combining this with Corollary~\ref{lem basis as perm}, we see that $\Delta_{I_i}$ is a sum over bijections between $I_i$ and $\mathcal{P}$.  The nonzero terms in this sum are precisely those bijections such that each propagator is associated to one of its supporting vertices in $I_i$, namely \hlfix{$I_i(p)$}{do we mean $I_i(p)$ or $V(p)$ here? if the first, then the last part of the sentence is wrong}, since only those locations in $C(W)$ are nonzero.  Since the nonzero entries of $C(W)$ are independent there can be no cancellation between terms in this expansion.

@@

All that remains is to prove that $\Delta_{I_i}$ has no irreducible factors of degree 3 or more.  Suppose for a contradiction that $f$ is a factor of $\Delta_{I_i}$ of degree $\geq 3$. Without loss of generality\footnote{This is without loss of generality for the following reason.  By removing the propagators which come before those contributing to $f$ (in the order given by $I_i$) and changing $i$ to be the first vertex which contributes to $f$, we obtain an admissible diagram.  In this diagram, $f$ still divides $\Delta_{I_i}$ and we also have $i\in I_i$ and that 
$i$ contributes to $f$.  Thus this diagram can be used in place of $W$.} we may assume $i\in I_i$ and that $i$ contributes to $f$.
Finally, we can suppose that $W$ is minimal in number of propagators with an irreducible factor of degree $\geq 3$.

Let $p$ be the propagator such that $I_i(p) = i$. There are two cases to consider: when $p$ is supported on the edge $i-1$ or the edge $i$.  These are illustrated in Figure~\ref{fig no big factors}.

@@

\textbf{Case 1}: Suppose $p = (i-1, m)$. Then $p$ has one end on the $i-1^{th}$ edge, and $I_{i+1}(p) = m$ by Remark~\ref{rmk cyclic}.

@@

If neither $m$ nor $m+1$ appear in $I_i$ then by Corollary~\ref{no coloops} we have $V(p) \cap I_i = \{i\}$, and so the row of $p$ in the matrix of $\Delta_{I_i}$ has only one nonzero entry; hence $\Delta_{I_i}$ has a linear factor contributed by $p$ and $i$, which is a contradiction to the fact that $i$ contributes to $f$.  So we must have at least one of $m$ and $m+1$ in $I_i$. However, all propagators in $S$ contribute to $I_{i}$ strictly before $m$ by \st{the previous paragraph}{\color{violet}Remark {clockwise ordering rem}}\note{The purple edit isn't true: a priori a propagator inside $p$ could lie on edge $m$ and contribute $m$. The previous paragraph explains why this doesn't happen.} so, after permuting rows and columns as needed, the matrix giving $\Delta_{I_i}$ has the form
\[
\begin{bmatrix} A & B \\ 0 & C\end{bmatrix}
\]
where $A$ is the $|S|\times |S|$ matrix indexed by the propagators in $S$ and the vertices in $I_i(S)$. No other propagators can be supported on these vertices since all other propagators are outside of $p$, and $p$ is the first propagator supported at $i$; this explains the zero block.  Therefore $\Delta_{I_i} = \det A \det C$, and both factors are nontrivial since at least one of $m$ and $m+1$ appear in $I_i$.  Since $p\in \cP_{in}(p)$ we have $f|\det A${\color{violet}. If we remove the propagators that contribute variables to the minor $C$,}\st{ and if we remove the propagator outside of $p$ that contributes $m$ or $m+1$,} we get a diagram $W'$ with fewer propagators for which $\det A | \Delta_{I^{W'}_i}$ still holds, hence $f|\Delta_{I^{W'}_i}$.  This contradicts the minimality of our choices. \sanote{trying to make clear that the minor changes with the diagram. May have just made it too notation heavy.}

\textbf{Case 2}: Suppose $p = (i, m)$, i.e. $p$ has an end on the $i^{th}$ edge. If no other propagators are supported on the vertex $i$ then the corresponding column of $C(W)$ has only one nonzero entry in it, and so $\Delta_{I_i}$ has a linear factor contributed by $p$ and $i$; as above, this is a contradiction. \st{Since $I_i(p) = i$ there can be no propagator supported on $i$ with an end on the edge $(i-1, i)$.  Thus to have another propagator supported on $i$ there is at least one other propagator on edge $(i,i+1)$.  Further since since $I_i(p)=i$, one of these other propagators contributes $i+1$ in $I_i$, so we can take $q$ to be the propagator such that $I_i(q)=i+1$.  We know that $q$ has one end on the edge $(i, i+1)$ and is adjacent to $p$ on that edge in the counterclockwise direction (see Figure {\ref{fig no big factors}}).}  \hlfix{Since $p$ is the first propagator in the ordering imposed by $I_i$, it is the counterclockwise most propagator supported on $i$ (by Remark {\ref{clockwise ordering rem}}). Since there are multiple propagators supported on $i$, there must be a $q$ ending on edge $i$, such that $I_i(q) = i+1$, that is second clockwisemost propagators supported by $i$.}{Alt. proof} Write $q = (i, j)$ for the support of $q$.  The situation for $q$ is very similar to case 1: \st{in particular, we have $I_{i+1}(q) = j$ by Lemma {\ref{vertex cyclic int lem}} and so if $j\in I_i$ then the propagator which contributes $j$ is outside of $q$.}{\color{violet}if we have $r$ such that $I_i(r) = j$, as above, $r$ lies outside $q$: $r \in \cP_{out}(q, <_i)$. }

Similarly to Case 1, let $S = \cP_{in}(q) \cup \{p\}$ in the $<_i$ order. Then all propagators in $S$ contribute to $I_i$ strictly before $j$ and no other propagators are supported on vertices strictly before $j$.  Thus the matrix giving $\Delta_{I_i}$ has the form {\color{violet}(when reordered according to how they contribute to $I_i$)}
\[
\begin{bmatrix} A & B \\ 0 & C\end{bmatrix}
\]
where $A$ is the submatrix indexed by the propagators in $S$ and the vertices in $I_i(S)$. Again two things can now happen.  If some vertex $j$ or larger (with respect to $<_i$) belongs to $I_i$ then $B$ and $C$ are at least one column wide, and so the block form of the matrix gives a nontrivial factorization of $\Delta_{I_i}=\det A\det B$.  This yields a contradiction as in Case 1: $f|\det A$ and removing a propagator not in $S$ gives a smaller diagram with a factor of degree $\geq 3$, which contradicts our minimality assumption.\note{I rephrased this last sentence quite a bit because earlier things had gotten rephrased} \sanote{reads fine to me.}

On the other hand, if no vertex $\geq_i j$ is in $I_i$ then we have $\Delta_{I_i} = \det A$.  Looking in more detail into $A$, note that the only vertices that support both $p$ and $q$ and also belong to $I_i$ are $i$ and $i+1$. Hence
\[
A = \begin{bmatrix} D & 0 \\ E & F\end{bmatrix}
\]
where $D$ is the $2\times 2$ matrix indexed by the propagators $p$ and $q$ and the vertices $i$ and $i+1$.  Thus $p$ and $i$ contribute to a quadratic factor of $\Delta_{I_i}$, once again contradicting our assumptions.

@@

\item \sanote{changed $b$ to $a+1$ here.} Suppose propagators $p$ and $q$ are supported on a common edge  $a$, with $I_i(p)=a$ and $I_i(q)=a+1$.  Let $x_{p,a},x_{p,a+1},x_{q,a},x_{q,a+1}$ be the associated variables in $C(W)$. For any fixed bijection $\sigma$ from $\cP\setminus\{p,q\}$ to $I_i \setminus\{a,a+1\}$ for which each propagator is supported on its image under the bijection, we can extend $\sigma$ to a bijection from propagators to $I_i$ in two ways: either $p\mapsto a$ and $q\mapsto a+1$ or $p\mapsto a+1$ and $q\mapsto a$.  The sum of the contributions of all these bijections to $\Delta_{I_i}$ is therefore the product of $x_{p,a}x_{q,a+1}-x_{p,a+1}x_{q,a}$ {\color{violet} By the arguments in the proof of the claim in part 2 of this theorem} \note{Could this use a ref back to where we discuss this kind of arugment in more detail.}\sanote{see violet?} with the minor coming from $\cP\setminus\{p, q\}$ and $I_i \setminus \{a,a+1\}$.  Since there is no cancellation of terms in the expansion of $\Delta_{I_i}$, if any other terms appear then they will cause a factor which is not in the form described in the previous part.  Therefore no such terms exist and $x_{p,a}x_{q,a+1}-x_{p,a+1}x_{q,a}$ is a factor of $\Delta_{I_i}$.

Now let $f$ be a quadratic factor of $\Delta_{I_i}$.  By part (2) we know that $f$ is a $2\times 2$ minor coming from two propagators, call them $p$ and $q$, and two vertices, call them $a <_i b$.  It remains to show that \st{$a$ and $b$ are adjacent} {\color{violet}$b = a+1$}.  From this we can conclude that $p$ and $q$ each have one end on edge $a$. \st{, as any other way for both $p$ and $q$ to be supported on two consecutive vertices would contradict noncrossing or the density requirement of admissibility.}

@@

\item If $S_i = \emptyset$ then $r_i = 1$ and we immediately have $r_i | \Delta_{I_i}$; otherwise, let $p\in S_i$. Note that $\Delta_{I_i}$ is homogeneous linear in the variables of the row corresponding to $p$.  By part (2), either exactly one variable in the row corresponding to $p$ appears in $\Delta_{I_i}$ and this variable is a factor of $\Delta_{I_i}$, or exactly two variables from the row corresponding to $p$ appear in $\Delta_{I_i}$ and they appear as part of a quadratic factor.  In the first case let the variable be $x$. Then $x$ is a factor of $\Delta_{I_i}$. {\color{violet} Therefore, $x$ must correspond to entry in the row of $p$ and column of $I_i(p)$, i.e. $x = x_{p, I_i(p)}$.} \st{and so in particular the monomial in $\Delta_{I_i}$ corresponding to the bijection between propagators and vertices of $I_i$ associates the column of $x$ to $p$. Thus $x$ also appears in $r_i$ and s} Since the matrix for $r_i$ is a minor of the matrix for $\Delta_{I_i}$ \st{and every term in $\Delta_{I_i}$ involves $x$,} we also have that \st{every term in $r_i$ involves $x$ so}{\color{violet}$x$ is a factor of $r_i$.} \st{$x$ is a factor of both $r_i$ and $\Delta_{I_i}$ and is the only variable from this row in either polynomial.}

Now suppose two variables from the row $p$ appear in a quadratic factor $f$ of $\Delta_{I_i}$.  By part (3), there is another propagator $q$ and an edge $a$ such that $f$ is the $2\times 2$ minor coming from $p, q$ and $a, a+1$, with \st{$I_i(p)=a$, $I_i(q)=b$}.  {\color{violet}There are two situations which can occur, either $I_i(p)=a$ and $I_i(q)=a+1$, or $I_i(q)=a$ and $I_i(p)=a+1$. These are} illustrated in Figure~\ref{fig quadratic}; we show that in both cases it follows that $q \in S_i$ as well.

\begin{figure}
\[
\begin{tikzpicture}[baseline=(current bounding box.east)]
	\begin{scope}
	\drawWLDfragment[2]{4}{0.1} 
	%\drawnumberspartial % use this to debug
	\newnode[below]{1}{a}
	\newnode[below]{2}{a+1}
	\begin{scope}
	\clipcenterarc(0,0)(\zero - 1.5*\step:\zero + 4.5*\step:\radius)
	\newprop[midway,left]{1.33}{-7}{\footnotesize $p$}
	\newprop[midway,right]{1.66}{10}{\footnotesize $q$}
	\end{scope}
	\node at (\zero + 0.6*\step:\radius*0.7) {\footnotesize $p$};
	\node at (\zero + 2.4*\step:\radius*0.7) {\footnotesize $q$};
\end{scope}
\end{tikzpicture}
\begin{tikzpicture}[baseline=(current bounding box.east)]
	\begin{scope}
	\drawWLDfragment[2]{4}{0.1} 
	%\drawnumberspartial % use this to debug
	\newnode[below]{1}{a}
	\newnode[below]{2}{a+1}
	\begin{scope}
	\clipcenterarc(0,0)(\zero - 1.5*\step:\zero + 4.5*\step:\radius)
	\newprop[midway,left]{1.33}{-7}{\footnotesize $q$}
	\newprop[midway,right]{1.66}{10}{\footnotesize $p$}
	\end{scope}
	\node at (\zero + 0.6*\step:\radius*0.7) {\footnotesize $q$};
	\node at (\zero + 2.4*\step:\radius*0.7) {\footnotesize $p$};
\end{scope}
\end{tikzpicture}
\]
  \caption{The situations giving a quadratic factor with variables appearing in $r_i$. }\label{fig quadratic}
\end{figure}\note{I'm not sure we really need this figure? (Figure~\ref{fig quadratic})}
